Interactive Curvature Tensor Visualization on Digital Surfaces\thanks{This work has been mainly funded by XXXXXXX research grants.}

### 2015-10-21 latex/article.tex
section 2 of article

\newcommand{\ie}{\emph{i.e.} }
\newcommand{\eg}{\emph{e.g.} }
\newcommand{\wnlog}{w.l.o.g. }
\newcommand{\Zr}{\ensuremath{\mathbb{Z}[\rho]}}
\newcommand{\C}{\ensuremath{\mathbb{C}}}
\newcommand{\E}{\ensuremath{\mathcal{E}}}

@@

\sloppy In our context, we consider digital shapes (any subset of $\Z^d$) and
boundaries of digital shapes. We denote by $\DSh$ the Gauss
digitization of $\Shape$ in a $d-$dimensional grid with grid step $h$,
\emph{i.e.}  $\DSh = \Shape \cap (h\Z)^d$. For such digitized set
$\DigShape$, $\Body{\DigShape}{h}$ \todo{DC: $\partial\Body{.}{h}$,
  pas $\Body{.}{h}$, c'est pas un bord fait de i-cell} denotes its
topological boundary, seen as a cellular Cartesian complex (See
Fig. \ref{fig:notations}-b). It is thus composed of $0$-cells and
$1$-cells (resp. \emph{pointels} and \emph{linels}), and, for $d=3$,
with $2$-cells ({\em surfels}), embedded in the digital grid.

\begin{figure}[t]{\small
    \begin{center}
      \subfigure[]{\begin{overpic}[width=5cm]{figs/integral2D_2}
          \put(42,28){$\times$}
          \put(46,31){$\vx$}
          \put(55,45){$B_R(\vx)$}
          \put(1,10){$\dS$}
          \put(20,10){$\Shape$}
      \end{overpic}}
      \subfigure[]{\begin{overpic}[width=5cm]{figs/notation_2}
          \put(44.5,25){$\vp$}
          \put(42,28.5){$\times$}
          \put(45,32.5){$\vx$}
          \put(45,16.5){$h$}
          %\put(88,39){$\Proj(p)$}
          \put(3,11){$\partial \Shape$}
          \put(15,7){$\Body{\DigShape}{h}$}
      \end{overpic}}
      \vspace{-0.4cm}
    \end{center}}
    \caption{Integral invariant computation $(a)$ and
      notations $(b)$ in dimension 2 \cite{CVIU2014}.\label{fig:notations}}
\end{figure}

% We define the volumetric integral as follow:
% \begin{Definition}
%   %
%   Given $\Shape \in \Shapes$ and a radius $R \in \R^{+*}$, the volumetric
%   integral $V_R(\vx)$ at $\vx \in \dS$ is given by:
%   %
%   \begin{equation}
%     V_R(\vx) \EqDef \int_{B_R(\vx)} \chi(\vp) d\vp \,,
%   \end{equation}
%   %
%   where $B_R(\vx)$ is the Euclidean ball of radius $R$ centered on $\vx$, and $\chi(\cdot)$ the characteristic function of $\Shape$. In dimension 2, we denote $A_R(\vx)$ such quantity.
%   %
% \end{Definition}

In \cite{DGCI2013}, we have defined the 2D digital curvature estimator
$\CurvH{R}$ and the 3D digital mean curvature estimator $\MeanCurvH{R}$ based on
the digital volumetric integral $V_{R_d}$ (or $A_{R_d}$ in dimension 2) of the
intersection between the shape $\Shape$ and a ball of digital radius $R_d$:

\begin{Definition}
  %
  Given $\DigShape \subset \Z^2$ (or $\Z^3$ for 3D mean curvature estimator) and
  $h$ a gridstep, {\em digital curvature estimators} are defined for any pointel
  $\vp \in \Body{\DigShape}{h}$ as:
  %
  \begin{align}
   \CurvH{R}(\DigShape,\vp) &\EqDef \frac{1}{h} \left( \frac{3 \pi}{2 R_d} - \frac{3 A_{R_d}(\vp)}{{R_d}^3} \right),\label{eq:dig-curvature-estimator-k}\\
   \MeanCurvH{R}(\DigShape,\vp) &\EqDef \frac{1}{h} \left( \frac{8}{3 R_d} - \frac{4V_{R_d}(\vp)}{{\pi R_d}^4} \right),\label{eq:dig-curvature-estimator-H}
  %  \\
  %  \hat{\kappa}^{1}(\DigShape,p) &\EqDef \frac{1}{h} \left( \frac{6(\hat{\lambda}_2 - 3\hat{\lambda}_1)}{\pi {{R_d}}^6} + \frac{8}{5{{R_d}}} \right),\label{eq:dig-curvature-estimator-k1}\\
  %  \hat{\kappa}^{2}(\DigShape,p) &\EqDef \frac{1}{h} \left( \frac{6(\hat{\lambda}_1 - 3\hat{\lambda}_2)}{\pi {{R_d}}^6} + \frac{8}{5{{R_d}}} \right),
  %  \label{eq:dig-curvature-estimator-k2}
  \end{align}
  %
  where $R_d = \frac{R}{h}$ is the digitized radius of the ball, and
  $A_{R_d}(p)$ and $V_{R_d}(p)$ are the number of digital points in the
  intersection between $\DigShape$ and the ball, $\DigShape\cap B_{R_d}(\vp)$.
  % Similarly, $\hat{\lambda}_1$
  % and $\hat{\lambda}_2$ are the two greatest eigenvalues of the covariance
  % matrix of the digital points in the intersection between $Z$ and the ball.
  %
\end{Definition}

We have proved that estimating the 2D curvature and 3D mean curvature on digital
objets with volumetric integral information have the multigrid convergence
property (setting $R = kh^\frac{1}{3}$ on convex shapes with at least
$C^3$-boundary and bounded curvature) \cite{CVIU2014}. For short, when the digital objet becomes
finer and finer, \ie when the digitization step $h$ tends to zero, the estimated
quantity on $\DSh$  converges (theoretically and experimentally) to
the associated one on $\partial\Shape$. More precisely, we have:

\begin{align}
  \CurvH{R}(\DSh,\vx,h) &= \Curv(\Shape,\vx) + O(h^{\frac{1}{3}}),\\
  \MeanCurvH{R}(\DSh,\vx,h) &= \MeanCurv(\Shape,\vx) + O(h^{\frac{1}{3}}) \,.
\end{align}

In \cite{CVIU2014}, we have also defined the 3D digital principal curvature estimators $\kappa^1_{R}$ and $\kappa^2_{R}$
on $Z\subset\Z^3$:
\begin{Definition}
  %
  Given $\DigShape \subset \Z^3$ and
  $h$ a gridstep, {\em 3D digital principal curvature estimators} are defined for any pointel
  $\vp \in \Body{\DigShape}{h}$ as:
  %
  \begin{align}
    \hat{\kappa}^{1}(\DigShape,p) &\EqDef \frac{1}{h} \left( \frac{6(\hat{\lambda}_2 - 3\hat{\lambda}_1)}{\pi {{R_d}}^6} + \frac{8}{5{{R_d}}} \right),\label{eq:dig-curvature-estimator-k1}\\
    \hat{\kappa}^{2}(\DigShape,p) &\EqDef \frac{1}{h} \left( \frac{6(\hat{\lambda}_1 - 3\hat{\lambda}_2)}{\pi {{R_d}}^6} + \frac{8}{5{{R_d}}} \right),
    \label{eq:dig-curvature-estimator-k2}
  \end{align}
  %
  where $R_d = \frac{R}{h}$ is the digitized radius of the ball, and
  $\hat{\lambda}_1$ and $\hat{\lambda}_2$ are the two greatest eigenvalues of the covariance matrix of the digital points in the intersection between $Z$ and the ball.
  %
\end{Definition}

% When dealing with digital data in a multigrid framework, we must pay
% attention to on how quantities expend where decreasing the grid step
% $h$. For example, if you fixed $R$, the digital radius will be $R_d
% = \frac{5}{h}$ and increase when $h$ tends to zero.
Authors of \cite{CVIU2014} have proved the convergence of these
estimators when setting the ball radius $h$ in $R = kh^\frac{1}{3}$,
where $k$ is a constant related to the maximal curvature of the
shape. More formally, we get:
\begin{align}
  \hat{\kappa}^1_{R}(\DigF{\Shape}{h},p) &= \kappa^1(X,x) + O(h^{\frac{1}{3}}),\\
  \hat{\kappa}^2_{R}(\DigF{\Shape}{h},p) &= \kappa^2(X,x) + O(h^{\frac{1}{3}}),
\end{align}
when the ball radius is $R = kh^\frac{1}{3}$.
Additionally, eigenvectors associated with $\hat{\lambda}_1$ and
$\hat{\lambda}_2$ correspond to principal curvature directions. The
smallest eigvenvector corresponds to the normal direction at $p$.
\todo[inline]{finir en citant le chapitre (ou CVIU) pour convergence directions}

\todo[inline]{petite phrase pour dire que R depend de la géométrie
  (citer DGCI pour paramter free, ou SMI pour dire qu'on peut s'en
  servir pour feature) et qu'il faut explorer R}

>>>>

\newcommand{\ie}{\emph{i.e.} }
\newcommand{\eg}{\emph{e.g.} }
\newcommand{\wrt}{\emph{w.r.t.} }
\newcommand{\wnlog}{w.l.o.g. }
\newcommand{\Zr}{\ensuremath{\mathbb{Z}[\rho]}}
\newcommand{\C}{\ensuremath{\mathbb{C}}}
\newcommand{\E}{\ensuremath{\mathcal{E}}}

@@

\sloppy In our context, we consider digital shapes (any subset of $\Z^d$) and
boundaries of digital shapes. We denote by $\DSh$ the Gauss
digitization of $\Shape$ in a $d-$dimensional grid with grid step $h$,
\emph{i.e.}  $\DSh = \Shape \cap (h\Z)^d$. For such digitized set
$\DigShape$, $\partial\Body{\DigShape}{h}$ denotes its
topological boundary, seen as a cellular Cartesian complex (See
Fig. \ref{fig:notations}-b). It is thus composed of $0$-cells and
$1$-cells (resp. \emph{pointels} and \emph{linels}), and, for $d=3$,
with $2$-cells ({\em surfels}), embedded in the digital grid.

\begin{figure}[t]{\small
    \begin{center}
      \subfigure[]{\begin{overpic}[width=5cm]{figs/integral2D_2}
          \put(42,28){$\times$}
          \put(46,31){$\vx$}
          \put(55,45){$B_R(\vx)$}
          \put(1,10){$\dS$}
          \put(20,10){$\Shape$}
      \end{overpic}}
      \subfigure[]{\begin{overpic}[width=5cm]{figs/notation_2}
          \put(44.5,25){$\vp$}
          \put(55,45){$B_R(\vx)$}
          \put(42,28.5){$\times$}
          \put(45,32.5){$\vx$}
          \put(45,16.5){$h$}
          %\put(88,39){$\Proj(p)$}
          \put(3,11){$\partial \Shape$}
          \put(15,7){$\partial \Body{\DigShape}{h}$}
      \end{overpic}}
      \vspace{-0.4cm}
    \end{center}}
    \caption{Integral invariant computation $(a)$ and
      notations $(b)$ in dimension 2 \cite{CVIU2014}.\label{fig:notations}}
\end{figure}

In \cite{DGCI2013}, we have defined the 2D digital curvature estimator
$\CurvH{R}$ and the 3D digital mean curvature estimator $\MeanCurvH{R}$ based on
the digital volumetric integral $V_{R_d}$ (or $A_{R_d}$ in dimension 2) of the
intersection between the shape $\Shape$ and a ball of digital radius $R_d$:
%
\begin{Definition}
  %
  Given the Gauss digitization $\Body{\DSh}{h}$ of a shape $\Shape \subset \R^2$
  (or $\R^3$ for the 3D mean curvature estimator) scaled by the digitization
  step $h$, {\em digital curvature estimators} are defined for any point $\vx
  \in \R^2$ (or $\R^3$) as:
  %
  \begin{align}
    \forall 0 < h < R,\quad
    &\CurvH{R}(\Body{\DSh}{h},\vx,h) \EqDef \frac{3\pi}{2R}
    - \frac{3\AreaC(\Ball{R/h}{\vx/h} \cap
    \Body{\DSh}{h}, h)}{R^3}\,,\label{eq:dig-curvature-estimator-k}\\
    &\MeanCurvH{R}(\Body{\DSh}{h},\vx,h) \EqDef \frac{8}{3R}
    - \frac{4\VolC(\Ball{R/h}{\vx/h} \cap
    \Body{\DSh}{h}, h)}{\pi R^4}\,.\label{eq:dig-curvature-estimator-H}
  \end{align}
  %
  where $\AreaC(\Ball{R/h}{\vx/h} \cap \Body{\DSh}{h}, h) = h^2 \MCard(
  \Ball{R/h}{\vx/h} \cap \Body{\DSh}{h} )$ and $\VolC(\Ball{R/h}{\vx/h} \cap
  \Body{\DSh}{h}, h) = h^3 \MCard( \Ball{R/h}{\vx/h} \cap \Body{\DSh}{h} )$ are
  the number of digital points in the intersection between $\Body{\DSh}{h}$ and
  the ball $\Ball{R/h}{\vx/h}$ of radius $R/h$ centered on $\vx/h$, scaled by
  $h^d$.\todo[inline]{J: Hum... pas top...}
  %
  % Similarly, $\hat{\lambda}_1$
  % and $\hat{\lambda}_2$ are the two greatest eigenvalues of the covariance
  % matrix of the digital points in the intersection between $Z$ and the ball.
  %
\end{Definition}
%
We have proved that estimating the 2D curvature and 3D mean curvature on digital
objets with volumetric integral information have the multigrid convergence
property (setting $R = kh^\frac{1}{3}$ on convex shapes, where $k$ is a constant
related to the maximal curvature of the shape, with at least $C^3$-boundary and
bounded curvature) \cite{DGCI2013}. For short, when the digital objet becomes
finer and finer, \ie when the digitization step $h$ tends to zero, the estimated
quantity on $\Bd{\Body{\DSh}{h}}$  converges (theoretically and experimentally)
to the associated one on $\Bd{\Shape}$. More precisely, we have:
%
\begin{align}
\forall 0 < h \le h_0,\,\, &\forall \vx \in \Bd{\Shape},\,\,
\forall \vxH \in \Bd{\Body{\DSh}{h}} \text{~with~} \| \vxH -\vx\|_\infty \le h, \nonumber \\
&\CurvH{R}(\DSh, \vxH, h) = \Curv(\Shape, \vx) + O\left(h^{\frac{1}{3}}\right)\,,\\
&\MeanCurvH{R}(\DSh, \vxH, h) = \MeanCurv(\Shape, \vx) + O\left(h^{\frac{1}{3}}\right)\,.
\end{align}

In \cite{CVIU2014}, we have also defined the 3D digital principal curvature
estimators $\PrincCurvH{1}{R}$ and $\PrincCurvH{2}{R}$ on $Z\subset\Z^3$ based
on digital moments:
%
\begin{Definition}
  %
  Given the Gauss digitization $\Body{\DSh}{h}$ of a shape $\Shape \subset \R^3$
  scaled by the digitization step $h$, {\em 3D digital principal curvature
  estimators} are defined for any point $\vx \in \R^3$ as:
  %
  \begin{align}
    \PrincCurvH{1}{R}(\Body{\DSh}{h},\vx,h) &\EqDef \frac{6}{\pi {R}^6}(\hat{\lambda}_2 - 3\hat{\lambda}_1) + \frac{8}{5{{R}}} \,,\label{eq:dig-curvature-estimator-k1}\\
    \PrincCurvH{2}{R}(\Body{\DSh}{h},\vx,h) &\EqDef \frac{6}{\pi {R}^6}(\hat{\lambda}_1 - 3\hat{\lambda}_2) + \frac{8}{5{{R}}} \,,
    \label{eq:dig-curvature-estimator-k2}
  \end{align}
  %
  where $R_d = \frac{R}{h}$ is the digitized radius of the ball, and
  $\hat{\lambda}_1$ and $\hat{\lambda}_2$ are the two greatest eigenvalues of
  the covariance matrix of the digital points in the intersection between
  $\Body{\DSh}{h}$ and the ball $\Ball{R/h}{\vx/h}$ of radius $R/h$ centered on
  $\vx/h$.\todo[inline]{J: Pareil ici... pas top...}
  %
\end{Definition}
%
The covariance matrix needs to compute digital moments of order $0$, $1$ and $2$
(see Equation~20 of \cite{CVIU2014} for more details). We have also proved the
convergence of these estimators when setting the ball radius $h$ in $R =
kh^\frac{1}{3}$, where $k$ is a constant related to the maximal curvature of the
shape, on convex shapes with at least $C^3$-boundary and bounded curvature
\cite{CVIU2014}. More formally, we get:
%
\begin{align}
  \forall 0 < h \le h_0,\,\, &\forall \vx \in \Bd{\Shape},\,\,
  \forall \vxH \in \Bd{\Body{\DSh}{h}} \text{~with~} \| \vxH -\vx\|_\infty \le h, \nonumber \\
  &\PrincCurvH{1}{R}(\DigF{\Shape}{h},\vxH, h) = \PrincCurv{1}(\Shape,\vx) + O\left(h^{\frac{1}{3}}\right) \,,\\
  &\PrincCurvH{2}{R}(\DigF{\Shape}{h},\vxH, h) = \PrincCurv{2}(\Shape,\vx) + O\left(h^{\frac{1}{3}}\right) \,.
\end{align}

Additionally, eigenvectors associated with $\hat{\lambda}_1$ and
$\hat{\lambda}_2$ of the covariance matrix correspond to principal curvature
directions $\PrincDirH{1}{R}$ and $\PrincDirH{2}{R}$. The smallest eigvenvector
corresponds to the normal direction $\NormalDirH{R}$ at $\vx$. Convergence
results can be found in \cite{ChapterIICurvature}.\todo[inline]{Changer
ChapterIICurvature dans le bibtex...}

We have shown that the radius of the ball depends to the geometry of the
underlying shape. In \cite{DGCI2014}, we proposed a parameter-free estimation of
the radius of the ball by analyzing the shape \wrt the local geometry through to
the maximal digital straight segments of the boundary of the digital shape. In
\cite{SMI2015}, we analyze in scale-space (\ie with a range of radii) curvature
estimation of our algorithm for a given digital shape. This allows us to detect
features of the shape thanks to the behavior on singularities. In both methods, we need to explore the radius of the ball at different sizes and the computation time, depending to the size of the digital objet and the radius of the ball, is an obstacle.
\todo[inline]{Faut faire une transition...}
Computing curvatures from our algorithm in real-time can be really useful.
